SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2018-09-28 scipy-1.0/paper.tex
Add links
Rearrange background
Write introduction

\begin{verbatim}
SciPy is not what you think it is

- At risk of failure many times
- Funding was *hard* to come by; funding agencies didn't easily give
  money.  Some labs, Enthough committed a bit of money.
- Java and other enterprize packages were getting support; MATLAB
  was huge.  Not many wanted to encourage this ``amateur'' effort.
- Graduate students & post-docs doing the heavy lifting; they found
  a system they enjoyed using, saw its potential, put energy into it.
- Driven by eclectic group of people being passionate about it, not
  because they were paid or told to work on it.
- And not by EECS people, but by scientists.  Some with fairly little
  programming background (David), others in varying stages of academic/research careers.
- Over time, the project became more more professional:
  - Standardized release formats
  - Docstrings
  - Testing & CI infrastructure
  - APIs
  - Issue tracking
  - Revision control (CVS -> SVN -> Git)
  - Conference
- Python wasn't perfectly suited in all ways; packaging, compiling,
  indexing (indexing/array protocols and complex numbers added)
- Initially, this high quality, stable, widely used project, hinged on
  three people doing the ``right'' thing to get it established.
- Bus factor for maintenance still very low.

The background section captures a selective history where we're trying
to identify some of the more important events.  But really, it relies
on the collective work of numerous individuals, too many to mention.
\end{verbatim}

@@

\subsection*{Conferences}

@@

\subsection*{Publications}
CiSE 2007 special issue

\subsection*{SciKits}

@@

\subsection*{Release History}

ANN: SciPy 0.10 -- Scientific Computing with Python

>>>>

%% SciPy is not what you think it is

%% - At risk of failure many times
%% - Funding was *hard* to come by; funding agencies didn't easily give
%%   money.  Some labs, Enthough committed a bit of money.
%% - Java and other enterprize packages were getting support; MATLAB
%%   was huge.  Not many wanted to encourage this ``amateur'' effort.
%% - Graduate students & post-docs doing the heavy lifting; they found
%%   a system they enjoyed using, saw its potential, put energy into it.
%% - Driven by eclectic group of people being passionate about it, not
%%   because they were paid or told to work on it.
%% - And not by EECS people, but by scientists.  Some with fairly little
%%   programming background (David), others in varying stages of academic/research careers.
%% - Over time, the project became more more professional:
%%   - Standardized release formats
%%   - Docstrings
%%   - Testing & CI infrastructure
%%   - APIs
%%   - Issue tracking
%%   - Revision control (CVS -> SVN -> Git)
%%   - Conference
%% - Python wasn't perfectly suited in all ways; packaging, compiling,
%%   indexing (indexing/array protocols and complex numbers added)
%% - Initially, this high quality, stable, widely used project, hinged on
%%   three people doing the ``right'' thing to get it established.
%% - Bus factor for maintenance still very low.

%% The background section captures a selective history where we're trying
%% to identify some of the more important events.  But really, it relies
%% on the collective work of numerous individuals, too many to mention.

SciPy recently released its first stable version: 1.0.  It signals to
the user community that the package's API has now reached a level of
maturity and stability that warrants trusting it inside research
pipelines.  But the version numbering belies the history of a package
that has already seen extensive adoption in research and industry.

Upon historic introspection, its arrival at this point is surprising,
and somewhat anomalous.  When first started, the library had very
little funding, and was written mainly by graduate students---many of
them without an official computer science education, or even the
blessing of their advisor.  To even imagine that a small group of
``rogue'' student programmers could upend the already well-established
research software ecosystem---backed by millions in funding, and
thousands of highly qualified engineers---was preposterous.

Yet, the philosophical motivations behind a fully open toolstack, and
an excited, friendly community with a singular focus, seemed to have
proven decisive in the long run.  It led not only to the library
described in this paper, but also to an entire ecosystem of related
packages, and a variety of social activities around them.  The
packages in the ecosystem share high standards of implementation,
documentation, and testing, and a culture eager to learn and adopt
better practices---both for community management and software
development.

Despite its success, SciPy still faces challenges: a low ``bus
factor'' (the minimum number of contributors that needs to leave to
cause serious difficult) being perhaps the most serious.

In the background section that follows, we capture a selective history
of some milestones and important events in the growth of SciPy.
Despite what we might highlight here, it is important to understand
that a project like SciPy is only possible because of the
contributions of very many contributors---too many to mention
individually, but each bringing an important part to the puzzle.

@@

\subsection*{SciPy Matures}

%Seeking articles for special issue on Python and Science and Engineering
%Paul Dubois
%Wed Apr 12 13:00:04 EDT 2006
%https://mail.python.org/pipermail/numpy-discussion/2006-April/007487.html
%
%Python--Batteries Included
%https://www.computer.org/csdl/mags/cs/2007/03/c3007.pdf
%
%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=4160244
%
%Python: Batteries Included
%Paul F. Dubois
%
%Python for Scientific Computing
%Travis E. Oliphant
%
%IPython: A System for Interactive Scientific Computing
%Fernando Perez ; Brian E. Granger
%
%Computational Physics Education with Python
%Arnd Backer
%
%Python Unleashed on Systems Biology
%Christopher R. Myers ; Ryan N. Gutenkunst ; James P. Sethna
%
%Reaching for the Stars with Python
%Perry Greenfield
%
%A Python Module for Modeling and Control Design of Flexible Robots
%Ryan W. Krauss ; Wayne J. Book
%
%Python in Nanophotonics Research
%Peter Bienstman ; Lieven Vanholme ; Wim Bogaerts ; Pieter Dumon ; Peter Vandersteegen
%
%Using Python to Solve Partial Differential Equations
%Kent-Andre Mardal ; Ola Skavhaug ; Glenn T. Lines ; Gunnar A. Staff ; Asmund Odegard
%
%Analysis of Functional Magnetic Resonance Imaging in Python
%K. Jarrod Millman ; Matthew Brett
%
%Python for Internet GIS Applications
%Xuan Shi
%
%Python for Education: Computational Methods for Nonlinear Systems
%Christopher R. Myers ; James P. Sethna
%
%Matplotlib: A 2D Graphics Environment
%John D. Hunter
%
%-----
%
%https://ieeexplore.ieee.org/xpl/tocresult.jsp?isnumber=5725228&punumber=5992
%
%Python for Scientists and Engineers
%K. Jarrod Millman ; Michael Aivazis
%
%
%Python: An Ecosystem for Scientific Computing
%Fernando Perez ; Brian E. Granger ; John D. Hunter
%
%The NumPy Array: A Structure for Efficient Numerical Computation
%Stefan van der Walt ; S. Chris Colbert ; Gael Varoquaux
%
%Cython: The Best of Both Worlds
%Stefan Behnel ; Robert Bradshaw ; Craig Citro ; Lisandro Dalcin ; Dag Sverre Seljebotn ; Kurt Smith
%
%Mayavi: 3D Visualization of Scientific Data
%Prabhu Ramachandran ; Gael Varoquaux
%
%??? these were in the same issue
%
%web2py for Scientific Applications
%Massimo Di Pierro
%
%From Equations to Code: Automated Scientific Computing
%Andy R. Terrel
%https://www.researchgate.net/publication/260585856_From_Equations_to_Code_Automated_Scientific_Computing

CiSE 2007 special issue

@@

SciPy 1.0 was released ... see timeline ...

1.1 ...

1.2

Python 2 / 3?

%\subsection*{SciKits}

@@
